---
title: "Class11"
author: "Matthew Tomaneng"
date: "5/8/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Downlad CSV file from PDB database <http://www.rcsb.org/stats.summary>. Read this into Ra nd ddtermine fraction of X-ray structures

```{r}
pdb.stats <- read.csv("Data Export Summary.csv")
```

Let's Calculate something

```{r}
percent <- (pdb.stats$Total / sum(pdb.stats$Total)) *100
names(percent) <- pdb.stats$Experimental.Method
percent
```

#Working with Bio3D
Lets load the package
```{r}
#install.packages("bio3d")
library(bio3d)
```

List the function and try them out

```{r}
pdb <- read.pdb("1hsg")
pdb
```

Q8 Write out a new C-Alpha atom only PDB file for viewing in VMD
```{r}
#select only C-alpha carbons
inds <- atom.select(pdb, "calpha")
#Check if we got all the c-alpha carbons
inds
```

Check if we have what we want
```{r}
pdb$atom[inds$atom,]
```

Let's Write out the and trim the PDB file
```{r}
ca.pdb <- trim.pdb(pdb, inds=inds)
ca.pdb
write.pdb(ca.pdb, file = "calpha_only.pdb")
```

Downloaded MUSCLE onto computer
```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```

Now we will try to superimpose all the proteins together
```{r}
# Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)
# Print to screen a summary of the 'pdbs' object
pdbs
```

FInd the sequence
```{r}
aa <- get.seq("1ake_A")
aa
```

Search the PDB database
```{r, cache = TRUE}
b <- blast.pdb(aa)
```
Plot the graph
```{r}
plot(b)
```

allign structures
```{r}
pdbs <- pdbaln(files)
ids <- basename.pdb
```

```{r}
# Access the first 5 rows, and 8 columns
pdbs$ali[1:5, 1:8]
# Associated residues numbers
pdbs$resno[1:5, 1:8]
```

Find Gaps in the data and remove them
```{r}
# Inspect alignment for gaps
gaps <- gap.inspect(pdbs$ali)
# Indices of non-gap columns
head(gaps$f.inds)
# Access non-gap columns
pdbs$ali[, gaps$f.inds]
```

Now to use the dat
```{r}
# Calculate sequence identity
seqidentity(pdbs)
# Calculate RMSD
rmsd(pdbs)
```

Now to start clustering and analyzing data
```{r}
# Calculate RMSD
rd <- rmsd(pdbs)
# Clustering
hc <- hclust(as.dist(rd))
grps <- cutree(hc, k=3)
# Plot results as dendrogram
hclustplot(hc, k=3)
```

PCA 
```{r}
# Perform PCA
pc <- pca(pdbs)
plot(pc, col=grps)
```

Normal Analysis
```{r}
# NMA all structures
modes <- nma(pdbs)
plot(modes, pdbs, col=grps, spread=TRUE)
```

