---
title: "Class12"
author: "Matthew Tomaneng"
date: "5/10/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Section 1
```{r}
#Load bio3d and download protein we're working with
#Already have it downloaded so just put it in the project enviornment
library(bio3d)
```
Next we will tranlate the file into something R can read
```{r}
hiv <- read.pdb("1hsg.pdb")
hiv
```
#Q1: What are the two non protein components
#A: Water and MK1 (ligand)

Section 1.2: Prepare initial protein and ligand input files
```{r}
#Trim data to components we can use isolated from each other
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")
#Now we are going to make these their own PDB files
write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, "1hsg_ligand.pdb")
```

Section 1.3: Using AutoDockTools to setup protein docking input
```{r}
#Most PDB files do not have charge and atom type so we need to add it using ADT

```

#Q2: Can you see the active site
#A: No, too convoluted with the structure and water atoms
```{r}
#Used ADT to add charges and hydrogen atoms to the PDB file and is saved as a pdbqt file
#We then made a grid box enclosing the active site
#Centered at (16,25,4) with dimmensions (30,30,30)
```

Section: 1.4 Prepare the ligand
```{r}
#Now doing the same thing done above but with the ligand file
```

Section: 1.5 Prepare a docking configuration file
```{r}
#Made a fonfig file for the dimmensions of our docking
#It will give the computer the info it needs to dock where we want it to
```

Section 2 Docking ligands into HIV-1 protease
```{r}
#Downloaded autdock vita to run the docking simulation
```
Section 2.2 Docking indinavir into HIV-1 protease
```{r}
# ~/Downloads/autodock_vina_1_1_2_mac/bin/vina --config config.txt --log
#log.tx
##This will run our docking simulation with the config file we made earlier

```

Section 2.3 Visualizing Results in R
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
#Go back to VMD to load the results
```

#Q4: Qualitatively how does the generated look to the actual
#A: Pretty darn close
```{r}
#Quantitative Analysis of the results
# res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("1hsg_ligand.pdbqt")
rmsd(ori, res)


```
#Q5: Quantitatively how good are the docks? Is the crystal binding mode reproduced within 1Å, RMSD for all atoms?
#A:Results are way above 1Angstrom so the results must not be too close

#Q6How would you determine the RMSD for heavy atoms only (i.e. non hydrogen atoms)?
#HINT: The atom.select() function will be of help here along with the selection string “noh” for no
#hydrogens.
```{r}
noh <- atom.select(read.pdb("1hsg_ligand.pdbqt"), type = "ATOM")
rmsd(noh, res)
```

