---
title: "Bioniformatics Class 7"
author: "Matthew Tomaneng"
date: "4/24/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Functions Again
Revisiting funtions from class 6
```{r}
source("http://tinyurl.com/rescale-R")
```

Lets see if we can use this function

```{r}
rescale(1:10)
```

Looks good lets break it!

```{r}
#rescale(c(1:10, "string"))
```

Lets try the new **rescale2** function

```{r}
#rescale2(c(1:10, "string"))
```

##Lets make a function to count number of NAs
Write a new function to check for NAs in two inputs

```{r}
#Start by searching for an existing function that does what you want
#example data
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```


```{r}
which(is.na(x))
```
How many total NAs
```{r}
sum(is.na(x) & is.na(y))
```

Let's test if this works for both vectors at once
```{r}
sum(is.na(c(x, y)))
```
#It works!
Now Let's Combine the Functions to make our function
```{r}
num_na <- function(x, y)
  #Number of NAs in the vectors
  sum(is.na(c(x, y)))
```
Let's test it this works

```{r}
num_na(x, y)
```

Test if it works for other situations
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
num_na(x, y2)
```

I messed up, doing it from both_na3
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
both_na3(x, y)
```

##Function for gene intersection
```{r}
df1
df2
x <- df1$IDs
y <- df2$IDs
```

```{r}
x
y
```

Found a function called Intersect
```{r}
intersect(x, y)
```
Not good enough, don't have position data
```{r}
#Value Matching funciton
x%in%y
```

```{r}
y%in%x
```
Getting a little more information when it comes to position of match
```{r}
#We can do better by getting the actual names of the genes
x[x%in%y]
```
Since X is a vector, putting the values into sqaure brackets will return the value

```{r}
#Now lets combine this info into a single table
#Put info into two columns 
cbind(x[x%in%y], y[y%in%x])

```

So now lets make this into a function

```{r}
#function(x, y) { 
#   cbind( x[ x %in% y ], y[ y %in% x ] )
#} Already made in the enviornment
#Not very clear as well
```

Lets try the merge function
```{r}
merge(df1, df2, by = "IDs")
```
##MERGE IS A SUPER HELPFUL FUNCTION













