---
title: "Bioinformatics Class 08"
author: "Matthew Tomaneng"
date: "4/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Unstructured learning: K clustering
```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

Our Tasks
Use the kmeans() function setting k to 2 and nstart=20
```{r}
km <- kmeans(x, centers = 2, nstart = 20)
print(km)
```


Inspect/print the results
Q. How many points are in each cluster?
```{r}
km$cluster
```
Q. What ‘component’ of your result object details
- cluster size?
```{r}
km$size
```

- cluster assignment/membership?
```{r}
km$cluster
```

- cluster center?
```{r}
km$centers
```
Plot x colored by the kmeans cluster assignment and
add cluster centers as blue points
```{r}
#Color Coding only the clusters so far
plot(x, col = km$cluster)
#Now we make the centers blue and with a different shape
points(km$centers, col = "blue", pch=15)
```
Repeat for k=3, which one has the better total SS? 
```{r}
#Repeat k cluster for k = 3
km2 <- kmeans(x, centers = 3, nstart = 20)
print(km2)
```

```{r}
plot(x, col = km2$cluster)
points(km2$centers, col = "blue", pch = 15)
```

#Hierarchial Clustering

Lets try on our X plot
```{r}
#First we need to calculate point(dis)similarity
#as the euclidean distance between observations
dist_matrix <- dist(x) 
hc <- hclust(d = dist_matrix)

# the print method is not so useful here so we use plot
hc
```
Lets draw the tree (plot)
```{r}
#You can choose how many clusters you want and simply move down #  from the top
plot(hc)
abline(h=6, col = "red")
cutree(hc, h=6) # Cut by height h
```

I can 'cut' the tree at any height to give our clustsers...

```{r}
#cut at height 6
cutree(hc, h = 6)
```

Or ask it to cut yieling 'k' clusters
```{r}
#give me 4 clusters
cutree(hc, k = 4)
```

Method of clustering matters too!
```{r}
#You can argue using: complete, single, average, centroid 
#hc.complete <- hclust(d, method="complete")
#hc.average <- hclust(d, method="average")
#hc.single <- hclust(d, method="single")
```


My Turn
```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters
```{r}
#ALWAYS NEED TO ADD DIST TO HCLUST
clust <- hclust(dist(x))
plot(clust)

```


```{r}
#to produce 2 clusters change out k
members2 <- cutree(clust, k = 2)
members3 <- cutree(clust, k = 3)

#Lets look at both (remember to cbind the new cutree info to be #  able to plot it)
cbind(k2 =members2, k3 = members3)
plot(x, col =members3, pch =15)
```

##How to do PCA in R

#prcomp function in PCA
Making up data
```{r}
## Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)
## Lets label the rows gene1, gene2 etc. to gene100
#Paste makes strings of character values
rownames(mydata) <- paste("gene", 1:100, sep="")
## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
## and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
 paste("ko", 1:5, sep="") )
## Fill in some fake read counts
for(i in 1:nrow(mydata)) {
 wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))

 mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
```

Use prcomp
```{r}
#prcomp flips data so you need to reflip it again (t = transpose = flips)
#head(t(mydata))
```

```{r}
pca <- prcomp( t(mydata), scale = TRUE )
pca
```

```{r}
## A basic PC1 vs PC2 plot 2-D
attributes(pca)
#x is  the name for our data so we use that in our PCA analysis
plot(pca$x[,1], pca$x[,2])

```

How well are our PC's capturing our data spread (i.e. variance)

```{r}
##percent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per
```

plot our scree plot

```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```

```{r}
#color up our PCA plot
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
```

Add some labels to points
```{r}
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
identify(pca$x[,1], pca$x[,2], labels=colnames(mydata))
```

Loading scores
```{r}
loading_scores <- pca$rotation[,1]
summary(loading_scores)
## We are interested in the magnitudes of both plus
## and minus contributing genes
gene_scores <- abs(loading_scores) 
```

UK Foods Hands On
```{r}
x <- read.csv("~/Downloads/UK_foods.csv")
dim(x)
```



















