---
title: "Class13"
author: "Matthew Tomaneng"
date: "5/15/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Genome Informatics (Part 1)
#Section 1
Q1: What are those 4 candidate SNPs? 

Q3: What is the location of rs8067378 and what are the different alleles for rs8067378? 
A:Chromosome 17: 39895095 (forward strand) A/G variants (43% G) 

Q4: What are the downstream genes for rs8067378? Any genes named ZPBP2,
GSDMB, and ORMDL3? 
A: GSDMB and ORMDL3 are downstream

Q5: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL)
are homozygous for the asthma associated SNP (G|G)?
[HINT: You can download a CVS file for this population from ENSEMBLE and use the R
functions read.csv(), and table() to answer this question] 
```{r}
mxl <- read.table("Mexican_genome.csv", header = TRUE)
table(mxl$X..Genotype)
#table gives us a summary of the different outputs for a column
nrow(mxl)
#number of total genomes present
9/64
# 14% is the answer
```

Q6. Back on the ENSEMBLE page, search for the particular sample HG00109. This is a
male from the GBR population group. What is the genotype for this sample?
A: G|G 


#Section 2: Initial RNA-Seq analysis 
Look more closely at sample HG00109 with G|G genotype for this SNP (from GBR population)
Q7: How many sequences are there in the first file?
What is the file size and format of the data? Make sure
the format is fastqsanger here!
A: 3863 sequences

Q8:  Does the first sequence have good quality? 
```{r}
#install.packages("seqinr")
#install.packages("gtools")
library(seqinr)
library(gtools)
phred <- asc(s2c("DDDDCDEDCDDDDBBDDDCC@")) - 33
phred

```
A: Yes decent quality

```{r}
#http://129.114.104.32/u/spongebob/h/class-13-matt
##This is a link to the history of using the supercomputer "Galaxy"
```

















