---
title: "Class16"
author: "Matthew Tomaneng"
date: "5/24/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#TnSeq Practice

Download important functions
```{r}
install.packages("tidyverse")
```

```{r}
#install.packages("tidyr")
#install.packages("dplyr")

library(tidyr)
library(dplyr)
library(tidyverse)
```

```{r}
library(ggplot2)
#install.packages("readr")
library(readr)
```

Loading file with experimental results
```{r}
exp_data <-read.csv("data/Experimental_results.csv")
head(exp_data)
```

Want to remove X column
```{r}
#exp_data <- exp_data[,-1]
#We can do it like this or we can use select
exp_data <- select(exp_data, -X)
head(exp_data)
```

```{r}
#Select function practice
mut_str_env <- select(exp_data, Mut_ID, Strain, Environment)
head(mut_str_env)
```

Let's get ready to start plotting
```{r}
#We need to make things easier to plot so we will make a column for frequency and a column for time
#We need to rotate the time values in the header to its own column, then rotate the frequency values as well
?gather
exp_rearranged<-gather(exp_data, Generation, Frequency,H_0:H_96)
head(exp_rearranged)
```

Now we are eliminating the "H" in the time data
```{r}
table_for_graph <- separate(exp_rearranged,Generation,into=c("H","Time"))
head(table_for_graph)
```

Next we want to get rid of the "H" from the graph entirely
```{r}
table_for_graph <- select(table_for_graph, -H)
head(table_for_graph)
```

Now we need to eliminate the NA's
```{r}
table_cleaned <- na.omit(table_for_graph)
table_cleaned$Time <- as.numeric(table_cleaned$Time)
head(table_cleaned)
```

Now we can finally plot our data
```{r}
#We can filter out by strain
DivAnc<-filter(table_cleaned, table_cleaned$Strain=="DivAnc")
L013<-filter(table_cleaned, table_cleaned$Strain=="L013")

```

```{r}
# make a plot for DivAnc strain
DivAnc_plot=ggplot(DivAnc)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#000033")+ggtitle("DivAnc_SC3")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
DivAnc_plot
```

```{r}
# make a plot for L013 strain
L013_plot=ggplot(L013)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#CC6633")+ggtitle("L013_SC3")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
L013_plot
```

We can combine these two graphs
```{r}
ggplot(table_cleaned)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#000033")+facet_grid(.~Strain)+ggtitle("Barcode trajectories")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
```

Now we can further analyze this data
```{r}
#We can look at specific mutations and their effects on frequency
# I've chosen Mut_ID==34
mut34<-filter(table_cleaned, table_cleaned$Mut_ID=="34")
mut34  
```

Now we plot the same thing but only for one mutation
```{r}
ggplot(mut34,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```

We see two clusters of data, so let's isolate one of the clusters
```{r}
mut34_f<-filter(mut34, mut34$Frequency>(-5))
mut34_f
```

Now we will plot the graph again but with the filtered data
```{r}
ggplot(mut34_f,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```

We have two outliers in the right graph that we need to get rid of
```{r}
ggplot(mut34_f,aes(Time, Frequency, colour = BC_ID, group=BC_ID))+geom_point()+geom_smooth(se = FALSE, method = "lm")+facet_grid(.~Strain)+theme(legend.position="none")+ggtitle(paste("Mutation",34, sep="_"))+xlab("Time, hours")+ ylab("Log10(Barcode frequency)")
```

Now it’s time to estimate slope for each barcode. Lets greate a file that will contain information about BC_ID, Mut_ID, Strain, and estimated slope.
```{r}
# Lets become familiar with lm function:

# For this exercise, take the filtered data for mutation 34 (mut34_f) and filter out information about one barcode you like

# I have chosen BC_ID=25361 in DivAnc strain
BC_25361<-filter(mut34_f, mut34_f$BC_ID=="25361", mut34_f$Strain=="DivAnc")
BC_25361
```

```{r}
#Lets plot frequency of this barcode:
ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")
```

```{r}
#Lets use lm function to fit the line to these points:
ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+geom_smooth(se = FALSE, method = "lm")+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")
```

Let's find the linear regression of this model to find out how accurate it is
```{r}
# Lets check what data does lm function return:
regression_model<-lm(Frequency~Time,BC_25361)
summary_data<-summary(regression_model)
summary_data
```

```{r}
# The information we are interested in is the value of Slopeand Intercept of this line:
# Let's try to access them:

# Time
Time<-summary_data$coefficients[2]
Time

```

```{r}
# Intercept:
Intercept<-summary_data$coefficients[1]
Intercept
```

Now we can find slopes for each barcode for each mutation in all strains.
```{r}
# Lets create the file:
data_header=matrix(data = NA,nrow = 1,ncol = 7)
        data_header[1]="Mut_ID"
        data_header[2]="BC_ID"
        data_header[3]="Strain"
        data_header[4]="Slope"
        data_header[5]="Intercept"
        data_header[6]="R^2"
write.table(data_header,"Data/Experimental_results.csv",append = FALSE, sep = ",",eol="\n",dec=".",row.names = FALSE,col.names = FALSE)
```






